\documentclass[defaultstyle,11pt]{thesis}

\usepackage{array}
\usepackage{epsfig}
\usepackage{url}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tabu}
\usepackage{tikz-qtree}
\usepackage{fancyvrb}

\usepackage{listings}
\lstset{
  language={},
  basicstyle=\footnotesize,
  numbers=left,
  numberstyle=\tiny,
  stepnumber=1,
  numbersep=5pt,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=4,
  captionpos=b,
  breaklines=true,
  breakatwhitespace=false,
  frame=single,
  frameround=tttt
}
\lstdefinelanguage{JavaScript}{
  keywords={break, case, catch, continue, debugger, default, delete,
    do, else, finally, for, function, if, in, instanceof, new, return,
    switch, this, throw, try, typeof, var, void, while, with, true,
    false, null},
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]',
  morestring=[b]'',
  sensitive=true
}

\clubpenalty = 10000
\widowpenalty = 10000

%%%%%%%%%%%%   All the preamble material:   %%%%%%%%%%%%

\title{Securing and Managing Trust in Modern Computing Applications}

\author{Andy}{Sayler}

\otherdegrees{B.S.E.E., Tufts University, 2011 \\
              M.S.C.S., University of Colorado, 2013}

\degree{Doctor of Philosophy}   %  #1 {long descr.}
	   {Dissertation Proposal}  %  #2 {short descr.}

\dept{Department of}		    %  #1 {designation}
     {Computer Science}         %  #2 {name}

\advisor{Prof.}                 %  #1 {title}
        {Dirk Grunwald}         %  #2 {name}

\reader{Prof. John Black}       %  2nd person to sign thesis
\readerThree{Prof. Sangtae Ha}  %  3rd person to sign thesis
\readerFour{Prof. Eric Keller}  %  4th person to sign thesis
\readerFive{Prof. Blake Reid}   %  5th person to sign thesis

\abstract{

  \OnePageChapter

  Today, users routinely push vast troves of private data into a
  myriad of cloud services, store personal information on mobile
  devices, and communicate about sensitive topics over public
  networks. These uses, while convenient, demonstrate a growing gap
  between users' desired use cases and their ability to control and
  protect their data. Today, users are often forced to choose between
  leveraging modern technologies and services, potentially exposing
  their data to a variety of third parties in the process, and
  maintaining privacy and control of their data.

  I feel these two choices represent a false dichotomy that can be
  resolved by disentangling the security and access control mechanisms
  used to protect user data from the underlying technologies and third
  parties that store and act on this data. In particular, I propose
  isolating the set of core user secrets (e.g. passwords,
  cryptographic keys, etc) that can be used to bootstrap the
  protection of all other user data and storing these core secrets
  with a dedicated, standardized, and service-oriented secret storage
  system. Such a Secret Storage as a Service (SSaaS) architecture
  allows users to reduce the trust they must place in individual
  technologies or third parties while still providing mechanisms to
  support the range of multi-user, multi-device use-cases popular
  today.

  In this dissertation proposal, I present an overview of the
  challenges to the security and privacy of user data inherent
  today. I introduce a framework for evaluating the degree to which we
  must trust various third parties inherent in a range of popular use
  cases, as well as the mechanisms by which this trust can be
  violated. I explore the problems with the traditional third party
  trust model and discuss ways in which SSaaS helps alleviate these
  issues. I discuss the qualities a desirable SSaaS ecosystem would
  have, present an SSaaS prototype, and propose several applications
  where the SSaaS model can significantly increase user security and
  privacy while also supporting a variety of modern use
  cases. Finally, I propose the additional research I will undertake
  to complete this dissertation, including a survey of modern services
  and technologies using the proposed trust-analysis framework, a more
  robust SSaaS survey implementation, and a variety of useful SSaaS
  client implementations.

}

%% \dedication[Dedication]{
%%   TBD
%% }

%% \acknowledgements{
%%   \OnePageChapter
%%   TBD
%% }

%\ToCisShort % use this only for 1-page Table of Contents

\LoFisShort	% use this only for 1-page Table of Figures
% \emptyLoF	% use this if there is no List of Figures

\LoTisShort	% use this only for 1-page Table of Tables
% \emptyLoT	% use this if there is no List of Tables

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%       BEGIN DOCUMENT...         %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\input macros.tex

\input chap_intro.tex
\input chap_background.tex
\input chap_challenges.tex
\input chap_related.tex
\input chap_trust.tex

%\input chap_ssaas.tex
\chapter{Secret Storage as a Service}
\label{chap:ssaas}

\section{Architecture}
\subsection{Stored Secrets}
\subsection{Secret Storage Providers}
\subsubsection{Secret Storage}
\subsubsection{Access Control}
\subsubsection{Auditing}
\subsection{Clients}
\section{Economics}
\section{Security and Trust}

%\input chap_custos.tex
\chapter{Custos: An SSaaS Prototype}
\label{chap:custos}

\section{Architecture}
\section{Protocol}
\section{Implementation}

%\input chap_apps.tex
\chapter{Applications and Use Cases}
\label{chap:apps}

\section{Storage}
\section{Communication}
\section{Authentication}

%\input chap_planned.tex
\chapter{Planned Work}
\label{chap:planned}

\section{Surveys}
\section{Multi-Party Sharding}
\section{Implementations}
\section{Analysis}

%\input chap_conclusion.tex
\chapter{Conclusion}
\label{chap:conclusion}


%%%%%%%%%   then the Bibliography, if any   %%%%%%%%%
\bibliographystyle{plain} % or "siam", or "alpha", etc.
\nocite{*}                % list all refs in database, cited or not
\bibliography{refs}       % Bib database in "refs.bib"

%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
\appendix
%\singlespacing
%\input appx_messages.tex
%\input appx_libcustos.tex

\end{document}

%%  LocalWords:  SSaaS tex Custos
